MCTS Based on Simple Regret

### 2012-02-22 slides/slides.tex
slides in progress

\title{Rational Monte Carlo Tree Search}
\author{David Tolpin, Solomon Eyal Shimony}
\institute{Ben-Gurion University of the Negev\\Beer Sheva, Israel}
\date{June 1, 2011}

@@

\begin{frame}{UCT}
UCT (Upper Confidence Bounds on Trees) is popular for Monte
  Carlo search in large trees:
\begin{itemize}
\item Selects an action $a_i$ that maximizes $\overline r_i + C\sqrt
  {\frac {\log n} {n_i}}$
\item Chooses a non-optimal action $n_i\ge\rho\log n$ times ($\rho$ is
  some constant).
\item Based on UCB1, that achieves logarithmic regret for multi-armed
  bandits.
\item But {\bf{\Large no bandits in Monte Carlo:}} {\bf no reward} is given for {\bf
  sampling} a good action.
\end{itemize}
\end{frame}

>>>>

\title{MCTS Based on Simple Regret}
\author{David Tolpin, Solomon Eyal Shimony}
\institute{Ben-Gurion University of the Negev\\Beer Sheva, Israel}
\date{February 22, 2012}

@@

\begin{frame}{Hard to solve search problems}
Search problems are often hard too solve in practice when:
\begin{itemize}
\item search space is extremely large;
\item {\it and} good heuristics are unknown.
\end{itemize}
\textbf{Easier} to solve:
\begin{itemize}
\item Chess --- search space size is manageable ($10^{50}$).
\item Timetabling --- good heuristics.
\end{itemize}
\textbf{Hard} to solve:
\begin{itemize}
\item Compute Go ($10^{180}$), Poker ($10^{70}$).
\item Canadian Traveller Problem.
\end{itemize}
\end{frame}

\begin{frame}{MCTS}

{\bf M}onte {\bf C}arlo {\bf T}ree {\bf S}earch helps in large search spaces.
\begin{itemize}
\item Starts with the root only.
\item Repeats:
  \begin{enumerate}
    \item \textbf{Selection:} select a branch to explore.
    \item \textbf{Expansion:} adds children of the leaf to the stored
      tree.
    \item \textbf{Simulation:} continues search (using a simple
      strategy) until a goal state is reached.
    \item \textbf{Backpropagation:} values of each stored node 
      are updated.
  \end{enumerate}
\end{itemize}
\textbf{Adaptive} MCTS samples `good' moves more frequently,\\
but sometimes {\bf explores} new directions.
\end{frame}

\begin{frame}{Multi-armed Bandit Problem and UCB}
Multi-armed Bandit Problem:
\begin{itemize}
\item We are given a set of $K$ arms.
\item Each arm can be pulled multiple times.
\item The reward is drawn from an {\bf unknown} (but normally {\it
    stationary} and {\it bounded}) distribution.
\item The {\bf total reward} must be maximized.
\end{itemize}

{\bf UCB} is near-optimal for MAB --- solves
{\it exploration/exploitation} tradeoff.
\begin{itemize}
\item pulls an arm that maximizes {\bf U}pper {\bf C}onfidence {\bf
    B}ound: $b_i=\overline X_i+\sqrt {\frac {c \log (n)} {n_i}}$
\item the cumulative regret is $O(\log n)$.
\end{itemize}
\end{frame}

\begin{frame}{UCT}
UCT ({\bf U}pper {\bf C}onfidence Bounds applied to {\bf T}rees) is 
based on UCB.
\begin{itemize}
\item Adaptive MCTS.
\item Applies the UCB selection scheme at each step of the rollout.
\item Demonstrated good performance in Computer Go (MoGo, CrazyStone, Fuego,
  Pachi, ...) as well as in other domains.
\end{itemize}
However, the first step of a rollout is different:
\begin{itemize}
\item The purpose of MCTS is to choose an action with the greatest utility.
\item Therefore, the {\bf simple regret} must be minimized.
\end{itemize}
\end{frame}

\begin{frame}{SRCR}

{\bf S}imple {\bf R}egret followed by {\bf C}umulative {\bf R}egret.
\begin{itemize}
\item Maximizes {\bf simple regret} at the {\bf first step}.
\item Continues with UCT from the {\bf second step on}.
\end{itemize}
\vspace{1em}
\begin{algorithmic}[1]
\Procedure{Rollout}{node, depth=1}
  \If {\Call{IsLeaf}{node, depth}}
    \State \textbf{return} 0
  \Else
    \State \textbf{if} depth=1 \textbf{then} action $\gets$ \Call{FirstAction}{node} \label{alg:srcr-first-action}   
    \State \textbf{else} action $\gets$ \Call{NextAction}{node} \label{alg:srcr-next-action}
    \State next-node $\gets$ \Call{NextState}{node, action}
    \State reward $\gets$ \Call{Reward}{node, action, next-node}
     \State \hspace{4em} + \Call{Rollout}{next-node, depth+1}
    \State \Call{UpdateStats}{node, action, reward}
  \EndIf
\EndProcedure
\end{algorithmic}
\end{frame}

\begin{frame}{Sampling for Simple Regret}
Sampling schemes for miniminizing the simple regret:
\begin{enumerate}
\item $\varepsilon$-greedy sampling.
\item a modified version of UCB (worse for cumulative, better for
  simple regret).
\item<+-> VOI-based sampling.
\end{enumerate}

\begin{itemize}
\item<+-> 1, 2 --- heuristic selection criterion, theoretical upper bounds
  can be obtained.
\item<+-> 3 --- based on principles of {\it Rational Metareasoning},
  but harder to analyze.
\end{itemize}

\end{frame}
